---
title: "Group 1 — Clinic bodies"
subtitle: "Studio wall"
author: "Vidhi, Sadiya, Madhumita, Dheemant"
format:
  html:
    toc: true
    toc-depth: 2
---

# Spine

> Who is well in the clinic, who is labelled ill, and where does infant death still sit on the map?
Both panels together show who in the clinic is labelled ill: how heart disease splits across sex and chest-pain type, and whether high blood sugar and exercise-induced angina go together more than chance would predict.


# Introduction — the tables we own

This folder is ours for the ten days. Three tables, three shapes.

- **Panel A** (`panel-A.csv`) — clinic measurements: age, blood pressure, cholesterol, maximum heart rate, plus sex and heart-disease label for grouping.
- **Panel B** (`panel-B.csv`) — the *same people*, Quals only: chest pain, blood sugar flag, exercise angina, heart disease.
- **Panel C** (`panel-C.csv`) — infant mortality by country. No geometry in the file; we join to a world map in R.

A does not grow a mosaic. B does not grow a honest histogram of chest pain. C is a rate, not a headcount.

We load from this folder:
```{r}
#| label: Set Up Chunk
library(tidyverse)
library(mosaic)
library(skimr)
library(vcd)
```

Found tables train the mark. A campus Free Hunch, if we collect one, is a **made** panel of the same pair type — new people, old grammar. It does not replace A, B, or C.


# Panel A — body / measure

*Age, BP, Cholesterol, MaximumHR are Quant. Sex and HeartDisease are Quals used to *split* the Quants.*

```{r}
#| label: My Panel A data
myPanelAData <-  read_csv(file = "data/panel-A.csv")
```


### 1. See the table

What is a case? Which columns are Quant / Qual? What is missing?

The Qualitative variable in this table entail - HeartDisease and Sex(These were character variables which were mutated to factor variables).

the quantitative variables in this table are - Age, BP, Cholestrol, MaximumHR.

A few piecies of data missing here are the fatality rate and the fact that casuality cannot be proved since this is not an RCT data set.

```r
# glimpse() / inspect() / skim()
```

```{r}
#| label: Table for Panel A
glimpse(myPanelAData)
skim(myPanelAData)
```

```{r}
#| label: change to factor(Panel A)
myPanelAData |> mutate(Sex = as.factor(Sex))
myPanelAData |> mutate(HeartDisease = as.factor(HeartDisease))
myPanelAMod <- myPanelAData |> dplyr::mutate(Sex = as.factor(Sex), HeartDisease = as.factor(HeartDisease))
glimpse(myPanelAMod)
```

### 2. Ask

One question *this* table can answer. One it cannot. Write both.

Question can answer:

1.How does cholestrol affect different sex?

2.Based on sex and age how does the maximum Heartrate fluctuate?

3.How does the age distribution of people with and without heart disease differ in this dataset?

4.how does the number of men and women having BP vary over the years?

Questions it cannot answer:

1.For what % of the population is cholestrol the reason for heart disease?

2.Does sex itself influence a person's maximum heart rate, or are the differences in maximum heart rate actually related to other factors such as age or health?

### 3. Choose a mark

 Questions can be answered:

```{r}
#| label: Chart1-Question1A
myPanelAData |> gf_boxplot(~ Cholesterol, bins = 50, fill = ~ Sex)
```

```{r}
#| label: Chart2-Question2A
myPanelAData |> gf_point(MaximumHR~Age, color = ~Sex) |>  gf_smooth(method = "lm") |>  gf_density_2d()
```

```{r}
#| label: Chart3-Question3A
myPanelAData |> gf_boxplot(Age ~ HeartDisease)
```

```{r}
#| label: Chart4 - Question4A
myPanelAMod |> gf_line(BP ~ Age,bins= 100, color =  ~Sex)
```

Questions that cannot be answered:
```{r}
#| label: Chart1- Question1AN
myPanelAData |> gf_point(Cholesterol~Sex,color = ~HeartDisease)
```

```{r}
#| label: Chart2-Question2AN
myPanelAData |> ggplot(aes(Sex, MaximumHR, fill = Sex)) + geom_violin(alpha = 0.6) + geom_jitter(width = 0.08, alpha = 0.3, size = 1) + theme_minimal()
```

```{r}
#| label: Statistic Test 1- Summarize
myPanelAData |> group_by(Sex) |> summarize( mean_MaximumHR = mean(MaximumHR, na.rm = TRUE))
myPanelAData |> group_by(Sex) |> summarize( mean_Cholesterol = mean(Cholesterol, na.rm = TRUE))
myPanelAData |> group_by(Sex) |> summarize( mean_BP = mean(BP, na.rm = TRUE))
```

```{r}
#| label: Statistic Test 2- T-Test
mosaic::t_test(MaximumHR ~ Sex, data = myPanelAData)
mosaic::t_test(Cholesterol ~ Sex, data = myPanelAData)
mosaic::t_test(BP ~ Sex, data = myPanelAData)

```

### 4. Predict: What do we think we will see?

We predict that age, sex, blood pressure, and cholesterol will show noticeable differences in maximum heart rate and heart disease.

### 5. Say

Title:
Age and sex show different patterns in heart health measures.

Caption:
The graphs show how age, sex, blood pressure, cholesterol, maximum heart rate, and heart disease are distributed and related across the sample.

One sentence we will stand behind:
The data shows differences between groups, but these patterns do not prove that one factor causes heart disease.

### 6. Limit
What this panel cannot do. Which kit table we refused to abuse.
The data available to us in this panel cannot prove dependency in terms of causality between two or more variables.



# Panel B — category / talk

*ChestPain, Sex, HeartDisease, ExerciseInducedAngina. Mosaic or stacked bar. No mean of ChestPain.*

```{r}
#| label: My Panel B Data
myPanelBData <-  read_csv(file = "data/panel-B.csv")
```

### 1. See the table

What is a case? Which columns are Quant / Qual? What is missing?

```r
# glimpse() / inspect() / skim()
```
```{r}
#| label: Table for Panel B
myPanelBData <-  read_csv(file = "data/panel-B.csv")
glimpse(myPanelBData)
inspect(myPanelBData)
skim(myPanelBData)
```

```{r}
#| label: Change to factor(Panel B)
myPanelBData |> mutate(Sex = as.factor(Sex))
myPanelBData |> mutate(HeartDisease = as.factor(HeartDisease))
myPanelBData |> mutate(ChestPain = as.factor(ChestPain))
myPanelBData |> mutate(ExerciseInducedAngina = as.factor(ExerciseInducedAngina))
myPanelBMod <- myPanelBData |> dplyr::mutate(Sex = as.factor(Sex), HeartDisease = as.factor(HeartDisease), ChestPain = as.factor(ChestPain), ExerciseInducedAngina = as.factor(ExerciseInducedAngina))
glimpse(myPanelBMod)

```

### 2. Ask

Questions that can be answered:

1. What part of the population posses both Blood Sugar and Heart Diseases?

2. How does the proportion of people with HeartDisease vary across different types of chest pain?

3. How does the distribution of Exercise Induced Angina differ between males and females? 

Questions that cannot be answered:

 1. How often does Exercise Induced Angina cause Chest Pain amongst men and women?
  
 2. Why do certain types of Chest Pain have higher rates of Heart Disease?
 
 3. What factors are responsible for certain people having heart diseases?
 
### 3. Choose a mark
Geom, map mark, or (later) test that matches the pair type. Name it before you code.

Questions that can be answered:
```{r}
#| label: Chart1-Question 1B

myPanelBData |> ggplot(aes(x = BloodSugar, fill = HeartDisease)) +geom_bar() +
  labs(x = "Blood Sugar", y = "Number of People")
```

```{r}
#| label: Chart2-Question2B
myPanelBData |> ggplot(aes(x = ChestPain, fill = HeartDisease)) + geom_bar(position = "fill") +labs(title = "Heart Disease Proportion Across Chest Pain Types",x = "Chest Pain Type",y = "Proportion",fill = "Heart Disease") + theme_classic()

```

```{r}
#| label: Chart3-Question3B
myPanelBData |> ggplot(aes(x = ExerciseInducedAngina, fill = HeartDisease)) + geom_bar(position = "dodge") +facet_wrap(~Sex) +
  labs(x = "Exercise-Induced Angina", y = "Number of People")
```

Questions that cannot be answered:
```{r}
#| label: Chart1-Question1BN
vcd::structable(ChestPain ~ ExerciseInducedAngina + Sex, data = myPanelBData) |> vcd::mosaic(labeling = labeling_values(), shade = TRUE, type = "expected")

```

```{r}
#| label: Chart2-Question 2BN
myPanelBData |> 
  ggplot(aes(x = ChestPain, fill = HeartDisease)) + geom_bar() + coord_flip() +
  labs(x = "Chest Pain", y = "Number of People")
```

```{r}
#| label: Chart3-Question3BN
vcd::structable(ExerciseInducedAngina ~ BloodSugar, data = myPanelBData) |> vcd::mosaic(labeling = labeling_values, shade = TRUE,gp = shading_Friendly, type = "expected")
```

Statistic
```{r}
#| label: Chi-Square Test
mosaic::xchisq.test(BloodSugar ~ ExerciseInducedAngina, data = myPanelBData)
mosaic::xchisq.test(ChestPain ~ ExerciseInducedAngina, data = myPanelBData)
```

Observations and Conclusions Drawn:

Blood Sugar ~ Exercise Induced Angina-
The above test proves that there is no significant correlation between the two variables. It states that the Blood Sugar and Exercise Induced Angina are independent of each other.
Further since the p value exceeds 0.5 it fails to prove the fact that nothing is happening.

Chest Pain ~ Exercise Induced Angina-
The above test proves that there is strong association between the two variables. Chest Pain and Exercise Induced Angina are Highly dependent on each other.
The low p value proves the hypothesis that nothing is happening. 

### 4. Predict: What do we think we will see?

We predict that heart disease will be more common among people with certain chest pain types, exercise-induced angina, and high blood sugar.

### 5. Say

Title:
Heart disease shows different patterns across several health-related factors.

Caption:
The graphs compare heart disease with chest pain, blood sugar, exercise-induced angina, and sex to identify patterns within the sample.

One sentence we will stand behind:
The data shows associations between heart disease and these factors, but it cannot tell us what causes heart disease.


### 6. Limit

What this panel cannot do:
This panel can show associations between the variables, but it cannot explain why these patterns occur or prove that one factor causes heart disease.

Which kit table we refused to abuse:
The contingency table.


# Panel C — place

*Join `country` to `rnaturalearth` / `tmap::World`. Fill is infant mortality (a rate).*
```{r}
#| label: My Panel C Data
myPanelCData <-  read_csv(file = "data/panel-C.csv")
```
Geometry is not in the CSV. Join in R, then use the same seven headings.

```r
# library(sf)
# library(rnaturalearth)
# world <- ne_countries(scale = "medium", returnclass = "sf")
# # join on country name; list the unmatched labels
```


### 1. See the table

What is a case? Which columns are Quant / Qual? What is missing?

```r
# glimpse() / inspect() / skim()
```


```{r}
myPanelCData <-  read_csv(file = "data/panel-C.csv")
glimpse(myPanelCData)
skim(myPanelCData)
```

### 2. Ask

One question *this* table can answer. One it cannot. Write both.

### 3. Choose a mark

Geom, map mark, or (later) test that matches the pair type. Name it before you code.

### 4. Predict

What do we think we will see? One sentence, before Run.

### 5. Make

Human brick first. Optional: one AI prompt, pasted below, only for what section 3 named.

```r
# human chunk

# ## Prompt we actually used
# ai_draft
# ours
```

### 6. Say

Caption + one sentence we will stand behind. Title is a sentence, not a column name.

### 7. Limit

What this panel cannot do. Which kit table we refused to abuse.


# Conclusion

What the spine looks like after three panels.

- The question we will defend in ninety seconds
- What we will **not** claim
- Which lines were human, which were machine
- Limits: sample, missingness, the test we did not run
- Gift for the steal-this strip (palette, annotation, join trick)

```r
# sessionInfo()
```
TODAY: Changes made in my file


